---
title: "Getting Started w/ Shiny"
author: "Jason Freels"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: slidy_presentation
css: shiny.css
runtime: shiny
---

# Overview

- Introduce Shiny and detail how it could benefit you

- Understand the different methods of building and sharing apps

- List some common steps for building shiny apps

- Build the framework for two common app architectures

    - `fluidPage`

    - `navbarPage`

- Walkthrough two apps to introduce reactivity 

# What is Shiny & Why Should I Use It?

## Shiny is...

- An R Package for Building Interactive Web Applications

- A direct link between R and various JavaScript libraries

## Why should you use Shiny?

- When I present a static plot (or when I’m presented a static plot) similar ‘what-if’ questions arise:

    1) What would happen to the plot (or result) if you changed the value of one or more model parameters

    2) What would happen if we chose a different statistical distribution?

    3) What would happen if we changed an important assumption?

-  Shiny allows us to show these alternate scenarios automatically

# There are multiple ways to build shiny apps

## Default rmarkdown

```{r echo=TRUE, eval=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Default Single File App

```{r echo=TRUE, eval=FALSE}
library(shiny)

# Define UI for application that draws a histogram
ui <- fluidPage(
   
   # Application title
   titlePanel("Old Faithful Geyser Data"),
   
   # Sidebar with a slider input for number of bins 
   sidebarLayout(
      sidebarPanel(
         sliderInput("bins",
                     "Number of bins:",
                     min = 1,
                     max = 50,
                     value = 30)
      ),
      
      # Show a plot of the generated distribution
      mainPanel(
         plotOutput("distPlot")
      )
   )
)

# Define server logic required to draw a histogram
server <- function(input, output) {
   
   output$distPlot <- renderPlot({
      # generate bins based on input$bins from ui.R
      x    <- faithful[, 2] 
      bins <- seq(min(x), max(x), length.out = input$bins + 1)
      
      # draw the histogram with the specified number of bins
      hist(x, breaks = bins, col = 'darkgray', border = 'white')
   })
}

# Run the application 
shinyApp(ui = ui, server = server)
```

## My Way - Works in all configuration `so far`

```{r echo=TRUE, eval=FALSE}
shinyApp(options = list(height = '600px'),

ui = fluidPage(theme = shinythemes::shinytheme('flatly'),'',
       sidebarLayout(
         sidebarPanel(
           selectInput("n_breaks", 
                       label = "Number of bins:",
                       choices = c(10, 20, 35, 50), 
                       selected = 20),
  
           sliderInput("bw_adjust", 
                       label = "Bandwidth adjustment:",
                       min = 0.2, 
                       max = 2, 
                       value = 1, 
                       step = 0.2)),
         mainPanel(plotOutput('test', height = '500px', width = '100%')))),
server = function(input, output, session) {
  
output$test <- renderPlot({
  
      hist(faithful$eruptions, 
           probability = TRUE, 
           breaks = as.numeric(input$n_breaks),
           xlab = "Duration (minutes)", 
           main = "Geyser eruption duration")
  
      dens <- density(faithful$eruptions, adjust = input$bw_adjust)
      
      lines(dens, col = "blue")
})
})
```

# How Will You Share Your App?

## Output formats

- As an HTML document

    + Provides a large canvas to discuss details of the app

    + Great for making interactive notes
    
- As an HTML~5~ presentation

    + Great for presenting teaching apps

    + Requires more focus on managing limited real-estate
    
- As an element on your website

    + Create interactive dashboards

    + Host instructional apps for your workforce

## Hosting services

- [Shiny Server](https://www.rstudio.com/products/shiny/shiny-server/) 

    - Standand alone, Linux-based infrastructure 

    - Provides a persistent connection between R serves and your app

    - Is free (except for Shiny Server Pro - $15K)

- [shinyapps.io](https://www.shinyapps.io)

    - Fully supported, Linux-based insfrastructure

    - Provides a persistent connection between R serves and your app

    - Has different pricing levels based on the amount of time your apps are running

# Helpful Shiny References

- [__shiny gallery__](http://shiny.rstudio.com/gallery)

- [__shiny user gallery__](https://www.rstudio.com/products/shiny/shiny-user-showcase/)

- [__show me shiny__](http://www.showmeshiny.com)

- [__shiny documentation__](https://cran.r-project.org/web/packages/shiny/index.html)

- [__R-bloggers__](http://www.r-bloggers.com/)

- [__shiny Google group__](https://groups.google.com/forum/#!forum/shiny-discuss/)

- [__StackOverflow__](http://www.stackoverflow.com/)

- [__HTML Widgets__](http://www.htmlwidgets.org)

- [__shiny GitHub page__](https://github.com/rstudio/shiny)

# Shiny App `navbar`

```{r}
library(shiny)
shinyApp(options = list(height = '800px', width = '100%'),
         
ui = navbarPage(theme = shinythemes::shinytheme('flatly'),
                includeCSS('../shiny.css'),
     tabPanel(h1('First Panel'),
        sidebarLayout(
          sidebarPanel(width = 3,
            
            selectInput("n_breaks1", 
                        label = h2("Number of bins:"),
                        choices = c(10, 20, 35, 50), 
                        selected = 20),
            
              sliderInput("bw_adjust1",
                        label = h3(HTML("<u>Bandwidth adjustment:</u>")),
                        min = 0.2,
                        max = 2,
                        value = 1,
                        step = 0.2)),
  
          mainPanel(width = 9,
                    plotOutput('jason', height = '500px', width = '100%')))),
  
    tabPanel(h4('Second Panel'),
       sidebarLayout(
         sidebarPanel(
           
            selectInput("n_breaks2", 
                        label = h5("Number of bins:"),
                        choices = c(10, 20, 35, 50), 
                        selected = 20),
  
            sliderInput("bw_adjust2", 
                        label = h6("Bandwidth adjustment:"),
                        min = 0.2, 
                        max = 2, 
                        value = 1, 
                        step = 0.2)),
         
        mainPanel(plotOutput('jason2', height = '500px', width = '100%'))))),

server = function(input, output, session) {
  
  output$jason <- renderPlot({
  
      hist(faithful$eruptions, 
           probability = TRUE, 
           breaks = as.numeric(input$n_breaks1),
           xlab = "Duration (minutes)", 
           main = "Geyser eruption duration")
  
      dens <- density(faithful$eruptions, adjust = input$bw_adjust1)
      
      lines(dens, col = "blue")
})
output$jason2 <- renderPlot({

      hist(faithful$eruptions, 
           probability = TRUE, 
           breaks = as.numeric(input$n_breaks2),
           xlab = "Duration (minutes)", 
           main = "Geyser eruption duration")
  
      dens <- density(faithful$eruptions, adjust = input$bw_adjust2)
      
      lines(dens, col = "blue")
})
})
```


